---
title: "Analyses with additional variables"
output: html_document
date: '2023-05-26'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)
library(lmerTest)
library(performance)
library(data.table)
library(dtplyr)
library(dplyr)
library(haven)
library(texreg)
library(writexl)
```

Grouping structure

1. Student pair
  - variables: Co-infection (outcome), Distance

2. Program
  - variables: Program size
  
3. School
  - variables: School size, Denomination, Avg income in postcode (~can affect multiple schools)
  
4. Gemeente
  - variables: 


# MLM overall

## Read data

```{r}
dat <- readRDS("H:/Mamba_3.10/additional_variables/net_bo.RDS")

# add program and school id to data
dat <- dat %>% 
  # also: recode coinfection
  mutate(across(matches("^coinfect_[0-9]"),
                ~case_when(.x == TRUE ~ 1,
                           .x == FALSE ~ 0,
                           is.na(.x) ~ 0))) %>% 
  group_by(WPOBRIN_crypt, WPOBRINVEST) %>% 
  mutate(school_id = cur_group_id()) %>% 
  ungroup() %>% 
  # (omit wpotypepo because all NA)
  group_by(WPOBRIN_crypt, WPOBRINVEST, WPOLEERJAAR, WPOOPLNR) %>%
  mutate(program_id = cur_group_id()) %>% 
  ungroup()

length(unique(dat$program_id))
length(unique(dat$school_id))
length(unique(dat$gemcode))

sum(dat$coinfect_14days)
sum(dat$coinfect_14days)/nrow(dat)
```

## Sampling

```{r}
# sample 5% of schools
set.seed(789)
school_sample <- dat %>% 
  select(school_id, school_size) %>% 
  distinct() %>% 
  slice_sample(weight_by = school_size, prop = 0.05)

dat_sch <- dat %>% 
  semi_join(y = school_sample)

# subset to complete cases from model with predictors
dat_sch <- 
  dat_sch[complete.cases(dat_sch[,colnames(m2c@frame)]),]

# add standardized predictors
dat_sch$distanceS <- scale(dat_sch$distance, scale = TRUE)
dat_sch$school_sizeS <- scale(dat_sch$school_size, scale = TRUE)
dat_sch$income_post_meanS <- scale(dat_sch$income_post_mean, scale = TRUE)

length(unique(dat_sch$program_id))
length(unique(dat_sch$school_id))
length(unique(dat_sch$gemcode))

skimr::skim(dat_sch)

with(dat_sch, summary(data.frame(distance, program_size, school_size, income_post_mean)))
                  
with(dat_sch, cor(data.frame(distance, program_size, school_size, income_post_mean), 
                  use = "complete.obs"))

table(dat_sch$WPODENOMINATIE)

sum(dat_sch$coinfect_14days)
sum(dat_sch$coinfect_14days)/nrow(dat_sch)
```

## Define Models

General formula: y ~ (1 | g1) + (1 | g1:g2) ...

  -> two hierarchical grouping factors (g1, g2), where g2 is nested in g1 (g1 would be schools and g2 classes e.g.)
  -> here: random intercept model, add variables between "(" and "|" for random slopes
  -> by default intercept and slope are correlated, to remove replace "|" with "||"
     
```{r}

# Singular fit, no variance in gemeente beyond program and schools
Sys.time()
m1_sch <- glmer(coinfect_14days ~ (1 | gemcode) + (1 | gemcode:school_id) + (1 | gemcode:school_id:program_id),
                family = binomial(), data = dat_sch)
Sys.time() #takes 50 mins

Sys.time()
m1_sch_nogem <- glmer(coinfect_14days ~ (1 | school_id) + (1 | school_id:program_id),
                      family = binomial(), data = dat_sch)
Sys.time() # takes 25 mins

Sys.time()
m1_sch_nopro <- glmer(coinfect_14days ~ (1 | gemcode) + (1 | gemcode:school_id),
                      family = binomial(), data = dat_sch)
Sys.time() # takes 18 mins

Sys.time()
m1_sch_gem <- glmer(coinfect_14days ~ (1 | gemcode),
                    family = binomial(), data = dat_sch)
Sys.time() # takes 6 mins

```

Add predictors

```{r}
# 2. Intercept only with predictors
Sys.time()
m2 <- glmer(coinfect_14days ~ 
              distance + program_size + school_size + WPODENOMINATIE + income_post_mean +
              (1 | school_id) + (1 | school_id:program_id),
            family = binomial(), data = dat_sch)
Sys.time()

summary(m2)


# With centered predictors
# also omit program size because wrongly calculated

dat_sch$distanceC <- scale(dat_sch$distance, scale = FALSE)
dat_sch$school_sizeC <- scale(dat_sch$school_size, scale = FALSE)
dat_sch$income_post_meanC <- scale(dat_sch$income_post_mean, scale = FALSE)
  
Sys.time()
m2b <- glmer(coinfect_14days ~ 
              distanceC + school_sizeC + WPODENOMINATIE + income_post_meanC +
              (1 | school_id) + (1 | school_id:program_id),
            family = binomial(), data = dat_sch)
Sys.time() # took 9hrs

summary(m2b)


# With centered AND SCALED predictors
# also omit program size because wrongly calculated
# and use gemeente and school, omit program
  
Sys.time()
m2c <- glmer(coinfect_14days ~ 
              distanceS + school_sizeS + WPODENOMINATIE + income_post_meanS +
              (1 | gemcode) + (1 | gemcode:school_id),
            family = binomial(), data = dat_sch)
Sys.time() # took 6hrs

summary(m2c)


# With centered AND SCALED predictors
# also omit program size because wrongly calculated
# and ONLY use gemeente, omit school and program

Sys.time()
m2d <- glmer(coinfect_14days ~ 
              distanceS + school_sizeS + WPODENOMINATIE + income_post_meanS +
              (1 | gemcode),
            family = binomial(), data = dat_sch)
Sys.time() # took 7hrs

summary(m2d)
```


# Final model export

```{r}
screenreg(list(m1_sch_nopro, m2c))

setwd("H:/Mamba_3.10/additional_variables/results")

# 1. Regression tables
# Output in word and later copy into excel
wordreg(m1_sch_nopro, file = "m1_null.doc")
wordreg(m2c, file = "m2_gemsch.doc")

# 2. Covariance matrix
vcov(m2c) %>% 
  as.matrix() %>% 
  as.data.frame() %>% 
  write_xlsx("H:/Mamba_3.10/additional_variables/results/m2_vcov.xlsx")
  
# 3a. Random effects gemeente
ranef(m2c, whichel = c("gemcode")) %>% 
  as.data.frame() %>% 
  write_xlsx("H:/Mamba_3.10/additional_variables/results/m2_ranf_gem.xlsx")

# 3b. Random effects school
ranef(m2c, whichel = c("gemcode:school_id")) %>% 
  as.data.frame() %>% 
  write_xlsx("H:/Mamba_3.10/additional_variables/results/m2_ranf_sch.xlsx")

# 4. Descriptive stats
skimdf <- dat_sch %>% 
  mutate(WPODENOMINATIE = as.factor(WPODENOMINATIE)) %>% 
  select(c('coinfect_14days', "distanceS", "school_sizeS", "WPODENOMINATIE", "income_post_meanS")) %>% 
  skimr::skim()
  
skimdf %>% 
  select(-numeric.hist) %>% 
  #modelsummary::datasummary_skim(histogram = FALSE) %>% 
  as.data.frame() %>% 
  write_xlsx("H:/Mamba_3.10/additional_variables/results/descriptives.xlsx")
```
